---
title: "Problem Set 2"
author: "Byeolha Kim"
date: "2025-02-12"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
Sys.setlocale("LC_ALL", "en_US.UTF-8")
```


## Reverse-engineer a table.

1. Using the dataset, recreate Table 1 of the Pickering and Davis article to the best of your ability. The numbers will not match exactly because you have different data. However, the table must have the same shape and include the same columns. Note that you may use the abbreviated country/module combo in place of the country names.

```{r create-table, echo=FALSE, message=FALSE, warning=FALSE}
load("C:/Users/Byeolha Kim/OneDrive - american.edu/SIS-750-004/PS2-main/ssa_water.Rdata")

# Create a summary table similar to Pickering and Davis (2012)
library(dplyr)
library(knitr)

# Define Mode function
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

summary_table <- water %>%
  group_by(countryModule) %>%
  summarise(
    Mean_Age = mean(age, na.rm = TRUE),
    Female_Ratio = mean(female == "Female", na.rm = TRUE),
    Mean_Household_Size = mean(householdSize, na.rm = TRUE),
    Mean_Education_Years = ifelse(is.nan(mean(eduYears, na.rm = TRUE)), 0, mean(eduYears, na.rm = TRUE)),
    Water_Access = Mode(waterSource),  
    Toilet_Type = Mode(toilet),        
    Wealth_Index = Mode(wealthIndex),  
    Rural_Percentage = mean(ruralHH == "rural", na.rm = TRUE),
    Sample_Size = n()
  )

# Print the table
kable(summary_table)
```

## Pay my respondents
2. Rather than sending a fully joined, final data frame, please copy and paste the output from a call to `str()` (e.g., `str(finaldat)`). 
```{r pay-respondents, echo=FALSE, message=FALSE, warning=FALSE}
# Load datasets
exp10 <- read.csv("C:/Users/Byeolha Kim/OneDrive - american.edu/SIS-750-004/PS2-main/exp10.csv")
exp11 <- read.csv("C:/Users/Byeolha Kim/OneDrive - american.edu/SIS-750-004/PS2-main/exp11.csv")
CRid <- read.csv("C:/Users/Byeolha Kim/OneDrive - american.edu/SIS-750-004/PS2-main/CRid.csv")

# Combine experimental data
exp_data <- bind_rows(exp10 %>% select(CompCode, Bonus), 
                      exp11 %>% select(CompCode, Pay)) %>%
  mutate(Bonus = coalesce(Bonus, Pay)) %>%
  select(-Pay)

# Merge with CRid data
finaldat <- inner_join(CRid, exp_data, by = "CompCode") %>%
  select(AssignmentID, rID, CompCode, Bonus)

# Output structure
kable(finaldat)
```
## Winning Wordle
3. Start by identifying the most common letters. Present a table (in descending order) of letters and their frequency of appearance in the dictionary. Hint: this isn't tidy data, and you'll have to tidy it first.
```{r wordle-analysis, message=FALSE, warning=FALSE}
# Load Wordle Dictionary
load("C:/Users/Byeolha Kim/OneDrive - american.edu/SIS-750-004/PS2-main/WordleDictionary.Rdata")

# Load required libraries
library(dplyr)
library(tidyr)
library(knitr)

# Count letter frequencies
letter_counts <- wordle %>%
  gather(position, letter, let1:let5) %>% 
  count(letter, sort = TRUE)

# Display letter frequencies
table_letters <- kable(letter_counts)

kable(table_letters)
```


4. Use the frequencies you just calculated to score each word (e.g., if "a" appears 10,000 times and "b" 450 times, then "babba" gets a score of 21,350). Report the highest and lowest scoring words and their scores.
```{r message=FALSE, warning=FALSE}
# Ensure letter_counts exists
if (!exists("letter_counts")) {
  letter_counts <- wordle %>%
    gather(position, letter, let1:let5) %>%
    count(letter, sort = TRUE)
}

# Compute word scores
word_scores <- wordle %>%
  rowwise() %>%
  mutate(score = sum(letter_counts$n[match(c(let1, let2, let3, let4, let5), letter_counts$letter)], na.rm = TRUE))

# Find the highest and lowest scoring words
highest <- word_scores %>% arrange(desc(score)) %>% head(1)
lowest <- word_scores %>% arrange(score) %>% head(1)

# Display highest and lowest scoring words
table_scores <- kable(rbind(highest, lowest))

print(table_scores)
```

5. Presumably I don't want to use words with duplicate letters. Among those with no duplicates, what are the best/worst opening words according to our standard?
```{r message=FALSE, warning=FALSE}
# Display highest and lowest scoring words
table_scores <- kable(rbind(highest, lowest))

# Find best and worst words with no duplicate letters
no_dups <- word_scores %>% filter(!dups)
best_no_dup <- no_dups %>% arrange(desc(score)) %>% head(1)
worst_no_dup <- no_dups %>% arrange(score) %>% head(1)

# Display best and worst words with no duplicates
table_no_dups <- kable(rbind(best_no_dup, worst_no_dup))

print(table_no_dups)
```

## Appendix: Replication Code

```{undefined message=FALSE, warning=FALSE}
# Appendix: Replication code

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
````